---
title: "Gene ordering"
author: "hamid"
date: '2023-01-13'
output: html_document
---

```{r}
# ORDERING GENE BASED ON CHROMOSOMAL LOCATION 
```{r}

# get gene location 

library(biomaRt)
library(BSgenome.Hsapiens.UCSC.hg19)
library(data.table)


mart <- useMart("ENSEMBL_MART_ENSEMBL", dataset = "hsapiens_gene_ensembl")

genes.table <- try(biomaRt::getBM(attributes = c("ensembl_gene_id", "external_gene_name",
"description", "gene_biotype", "chromosome_name", "start_position",'end_position','entrezgene_id'), mart = mart, useCache = F))  





x=fread('D:/DEEP_learning_project/df_transpose.csv')
x=as.data.frame(x)

# remove value before | of interest only 
colnames(x)[2:30] <- gsub(x = colnames(x)[2:30],pattern = "?|", replacement = "",fixed = TRUE)
names(x)[2:30] <- as.character(genes.table$external_gene_name[match(names(x)[2:30], genes.table$entrezgene_id)])

# remove everything after | ( included ) for each gene name
colnames(x)[31:length(colnames(x))]  <- gsub("\\|.*","",colnames(x)[31:length(colnames(x))])

# remove NA colnames who didn't match 
x <- x[!is.na(names(x))]



lol=x
get_gene_name=as.data.frame(colnames(lol))
colnames(get_gene_name)='gene_names'
get_gene_name=left_join(get_gene_name,genes.table, by = c("gene_names"="external_gene_name"))
get_gene_name=get_gene_name[!is.na(get_gene_name$start_position),]
get_gene_name=get_gene_name[!is.na(get_gene_name$end_position),]
get_gene_name=get_gene_name[!is.na(get_gene_name$chromosome_name),]
for (i in 1:length(get_gene_name$gene_names)){
  print(i)
  get_gene_name$mean_position[i]=(as.double((get_gene_name$end_position[i])+get_gene_name$start[i])/2)}

# order 
get_gene_name=get_gene_name[order( get_gene_name[,5], get_gene_name[,9] ),]

# get genes as vector 
avector=get_gene_name$gene_names
avector=unique(avector)

# drop columns in x for  gene with no chromosome location and no start/end site
x=x[avector]

# reorder on x 
x <- x[, avector]

# save
fwrite(x = x,file='D:/DEEP_learning_project/gene_data_ordered.csv',sep = ';')
```

